\bf Manifold Clustering for Latent Structure Block Models

### 2022-03-02 draft/draft.Rmd
updated draft

```{r, cache = TRUE}
maxit <- 50
animate <- TRUE
animation.dir <- '~/dev/manifold-block-models/draft'
random.starts <- 2
eps <- 1e-3
out <- lapply(seq(random.starts), function(i) {
  set.seed(i)
  intercept <- (i / random.starts > .5 + eps)
  manifold.clustering(Xhat, 2, 
                      parallel = TRUE, 
                      intercept = intercept, 
                      maxit = maxit,
                      eps = 1e-3,
                      verbose = FALSE,
                      animate = animate,
                      animation.dir = animation.dir,
                      animation.title = i)
})
names(out) <- paste('random', seq(random.starts))
names(out) <- sapply(seq(random.starts), function(i) {
  if (i / random.starts < .5 + eps) {
    paste('random', i, '\n(w/o intercept)')
  } else {
    paste('random', i, '\n(w/ intercept)')
  }
})

@@

\begin{example}

>>>>

```{r, cache = TRUE}
maxit <- 50
animate <- FALSE
animation.dir <- '~/dev/manifold-block-models/draft'
random.starts <- 2
eps <- 1e-3
out <- lapply(seq(random.starts), function(i) {
  set.seed(i)
  intercept <- (i / random.starts > .5 + eps)
  manifold.clustering(Xhat, 2, 
                      parallel = TRUE, 
                      intercept = intercept, 
                      maxit = maxit,
                      eps = 1e-3,
                      verbose = FALSE,
                      animate = animate,
                      animation.dir = animation.dir,
                      animation.title = i)
})
names(out) <- paste('random', seq(random.starts))
names(out) <- sapply(seq(random.starts), function(i) {
  if (i / random.starts < .5 + eps) {
    paste('random', i, '\n(w/o intercept)')
  } else {
    paste('random', i, '\n(w/ intercept)')
  }
})

@@

\begin{example}[Macaque visuotactile brain areas and connections]

```{r}
data(macaque, package = 'igraphdata')
A <- as.matrix(igraph::as_adjacency_matrix(macaque, type = 'both'))
A <- sign(A + t(A))
z <- as.numeric(factor(igraph::vertex_attr(macaque)$shape))

Xhat <- embedding(A, 2, 0)
plot(Xhat, col = z, asp = 1)
```

```{r cache = TRUE}
set.seed(123)
z.spectral <- spectral.clustering(A)
macaque.out <- manifold.clustering(Xhat, 2, 
                                   initialization = z.spectral, 
                                   intercept = TRUE, 
                                   parallel = TRUE, 
                                   verbose = FALSE, 
                                   animate = animate, 
                                   curve.init = 'random',
                                   animation.title = 'macaque')
```

```{r}
t.1 <- seq(min(macaque.out$t[[1]]), max(macaque.out$t[[1]]), .01)
t.2 <- seq(min(macaque.out$t[[2]]), max(macaque.out$t[[2]]), .01)
Y1 <- bezier.curve(t.1, macaque.out$p[[1]])
Y2 <- bezier.curve(t.2, macaque.out$p[[2]])

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], 
                 colour = factor(macaque.out$z)),
             size = .5) + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  coord_fixed() +
  theme(legend.position = 'none') +
  geom_path(aes(x = Y1[, 1], y = Y1[, 2])) +
  geom_path(aes(x = Y2[, 1], y = Y2[, 2])) + 
  xlim(min(Xhat[, 1]), max(Xhat[, 1])) + ylim(min(Xhat[, 2]), max(Xhat[, 2]))
```

\begin{example}[Non-intersecting curves]

```{r}
n1 <- 2 ** 7
n2 <- 2 ** 7
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
u1 <- runif(n1)
u2 <- runif(n2)
x1 <- cos(pi / 3 * u1)
y1 <- sin(pi / 3 * u1)
x2 <- 1 + cos(pi / 3 * u2 + pi)
y2 <- 1 + sin(pi / 3 * u2 + pi)
data.matrix <- cbind(c(x1, x2), c(y1, y2))

plot(data.matrix, col = z * 2, asp = 1, xlab = NA, ylab = NA,
     main = 'Latent configuration')
```

```{r}
P <- data.matrix %*% diag(c(1, 1)) %*% t(data.matrix)
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
plot(Xhat, col = z * 2, asp = 1, main = 'ASE')
```

```{r cache = TRUE}
z.init <- hclust(dist(Xhat), method = 'single') %>% 
  cutree(2)
nonintersect.out <- manifold.clustering(Xhat, 2, 
                                        initialization = z.init,
                                        intercept = TRUE, 
                                        animate = animate,
                                        animation.title = 'nonintersect')
```

```{r}
t.1 <- seq(min(nonintersect.out$t[[1]]), max(nonintersect.out$t[[1]]), .01)
t.2 <- seq(min(nonintersect.out$t[[2]]), max(nonintersect.out$t[[2]]), .01)
Y1 <- bezier.curve(t.1, nonintersect.out$p[[1]])
Y2 <- bezier.curve(t.2, nonintersect.out$p[[2]])

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], 
                 colour = factor(nonintersect.out$z)),
             size = .5) + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  coord_fixed() +
  theme(legend.position = 'none') +
  geom_path(aes(x = Y1[, 1], y = Y1[, 2])) +
  geom_path(aes(x = Y2[, 1], y = Y2[, 2])) + 
  xlim(min(Xhat[, 1]), max(Xhat[, 1])) + ylim(min(Xhat[, 2]), max(Xhat[, 2]))
```

### 2022-03-02 draft/draft.tex
update to draft
updated draft

>>>>

\begin{example}[Macaque visuotactile brain areas and connections]

\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-7-1} \end{center}

\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-9-1} \end{center}

\begin{example}[Non-intersecting curves]

\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-10-1} \end{center}

\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-11-1} \end{center}

\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-13-1} \end{center}

\hypertarget{simulation-study}{%
\section{Simulation Study}\label{simulation-study}}
